The Astropy Project

### 2017-11-20 main.tex
fixed typo
removed notes and todo

\section*{\textit{Notes and guidelines (to be removed)}}

\begin{itemize}
   \item The goal is to produce a brief and informative paper that covers major
        \astropy principles not mentioned in the first paper, the core v2.0
        package, and infrastructure in \astropy project to support development
        in \python.
    \item We don't plan on including code in this paper, but if you think you
        will need to include code in your section, please add it to a separate
        \python module (.py file) and include it in this repository.
    \item Use \escapecmd{\sectionname} not ``Section,'' \escapecmd{\figurename}
        not ``Figure''
    \item Use \escapecmd{\astropypkg} for the astropy package,
        \escapecmd{\astropy} for the astropy project, \escapecmd{\python} not
        ``Python''
    \item If your subpackage was included in Paper I, then please just include a
        note on what the package does, a reference to paper I, and any new major
        updates to your package
    \item If your subpackage was not included, then please describe the
        subpackage on level with what was in the first paper, and highlight any
        major features in it. Typical length should be equivalent to one page
        (single-column).
    \item Please make sure you are logged into Overleaf or pushing a commit with
        your information to be able to track the contributors to the paper.

\end{itemize}

\section*{Global TODO items}
\begin{itemize}
    \item \sout{subpackage not sub-package!} BMS: checked as done 13/11
    \item spell check!
    \item check consistent astropy usage
\end{itemize}

@@

\section{Astropy Core Package v2.0}
\label{sec:core}
% Adrian
The \astropy project aims to provide \python-based packages for all tasks that
are commonly needed in a large subset of the astronomical community.
At the foundation is the \astropypkg core package, which provides general
functionality (e.g., coordinate transformations, reading and writing astronomical
files) or base classes for other
packages to utilize for a common interface (e.g., \texttt{NDData}).
In this section, we highlight new features introduced or substantially improved
since version v0.2 (previously described in \citealt{astropy}).  The \astropkg 
provides a full log of changes \footnote{https://github.com/astropy/astropy/blob/master/CHANGES.rst} 
over the course of the entire project and more details about individual 
subpackages are available in the documentation \footnote{http://docs.astropy.org/en/stable/}.  
Beyond what is mentioned below, most subpackages have seen increased performance
since the release of the v0.2 package.

% \subsection{Analytic Functions}
% moved to models and depreciated --
% just here for completeness at the moment
%\inlinecomment{BMS}{I suggest to order the subsection lexicographically as reading the draft the current order seemed very random (unlike at the docs page where they the grouping logic is obvious)}
%\inlinecomment{Tom R}{the only issue is if we rely on having e.g. units described first if they are used later}

@@

The \texttt{astropy.units} subpackage adds support for representing units and
numbers with associated units --- ``quantities'' --- in code.
Historically, quantities in code have often been represented simply as numbers,
with units implied or noted via comments in the code because of considerations
about speed: having units associated with numbers inherently adds overhead to
numerical operations.
In \texttt{astropy.units}, \texttt{Quantity} objects extend \texttt{numpy}
array object and have been designed with speed in mind.

%\inlinecomment{Perry G}{Agreeing with Tom below, I would avoid using terms like subclass, decorators, mixin, namespace, (and perhaps even class, instead using objects consistently; this is no place to be pendantic about the distinction between classes and objects; even to the extent of adding a qualifying sentence somewhere like: "In this paper we will use object when we mean object or classes; in software there are important distinctions between the two terms, but they are not relevant in this paper."), and the like and replace them with more generic descriptions, e.g., "extended", or "tools to make defining functions that can use quantities simple to write" }
%\inlinecomment{Tom R}{I think saying 'subclass' here is too detailed and is mentioned below in 3.1.1 if needed - 'make it possible to attach units to numpy arrays' or something like that?}

@@

The \texttt{Quantity} object extends 
the \texttt{numpy.ndarray} object and therefore
works well with many of the functions in \texttt{numpy} that support
array operations. For example, \texttt{Quantity} objects with angular
units can be directly passed in to the trigonometric functions implemented in
\texttt{numpy}. The units are internally converted to radians (what the \texttt{numpy}
trigonometric functions expect) before being passed to \texttt{numpy}.

%\subsubsection{Speed improvements}

%\inlinecomment{Tom R}{I wonder if we might just want to mention at the top level that performance has been improved in places? This seems a bit too detailed}

%The \texttt{Quantity} object is now significantly
        %faster than in early versions of \astropypkg. This is in large part due
        %to the fact that \texttt{Quantity} subclasses the \texttt{numpy.ndarray}
        %object, but even relative to more recent versions of \astropypkg,
        %operations with \texttt{Quantity} objects are generally faster. This is
        %largely thanks to improvements in the way \texttt{numpy} arrays can be
        %subclassed.

@@

The \texttt{astropy.constants} subpackage provides a selection of physical and
astronomical constants as \texttt{Quantity} objects (see
\sectionname~\ref{sec:units}).
A brief description of this package was given in \citep{astropy}.
In version 2.0, the built-in constants have been organized into modules for
specific versions of the constant values.
For example, physical constants have \texttt{codata2014} \citep{codata2014} and
\texttt{codata2010} versions.
Astronomical constants are organized into \texttt{iau2015} and \texttt{iau2012}
modules to indicate their sources (resolutions from the International
Astronomical Union, IAU).
The \texttt{codata2014} and \texttt{iau2015} versions are combined into the
default constant value version: \texttt{astropyconst20}.
For compatibility with \astropypkg version 1.3,  \texttt{astropyconst13}
is available and provides access to the adopted versions of the
constants from earlier versions of \astropypkg.
To use previous versions of the constants as \emph{units} (e.g., solar masses),
the values have to be imported directly; with version
2.0, \texttt{astropy.units} uses the \texttt{astropyconst20} versions.

@@

New model classes can be created in two ways:
\begin{enumerate}
   \item The simplest way is to use the \texttt{custom\_model} decorator in modeling with a user-defined function that takes the inputs and model parameters as arguments.
   \item It is also possible to create an arbitrarily complex custom model baed on the \texttt{Model} class.
\end{enumerate}

>>>>

@@

\section{Astropy Core Package v2.0}
\label{sec:core}
% Adrian
The \astropy project aims to provide \python-based packages for all tasks that
are commonly needed in a large subset of the astronomical community.
At the foundation is the \astropypkg core package, which provides general
functionality (e.g., coordinate transformations, reading and writing astronomical
files) or base classes for other
packages to utilize for a common interface (e.g., \texttt{NDData}).
In this section, we highlight new features introduced or substantially improved
since version v0.2 (previously described in \citealt{astropy}).

% \subsection{Analytic Functions}
% moved to models and depreciated --
% just here for completeness at the moment
\inlinecomment{BMS}{I suggest to order the subsection lexicographically as reading the draft the current order seemed very random (unlike at the docs page where they the grouping logic is obvious)}
\inlinecomment{Tom R}{the only issue is if we rely on having e.g. units described first if they are used later}

@@

The \texttt{astropy.units} subpackage adds support for representing units and
numbers with associated units --- ``quantities'' --- in code.
Historically, quantities in code have often been represented simply as numbers,
with units implied or noted via comments in the code because of considerations
about speed: having units associated with numbers inherently adds overhead to
numerical operations.
In \texttt{astropy.units}, \texttt{Quantity} objects subclass the \texttt{numpy}
array object and have been designed with speed in mind.

\inlinecomment{Perry G}{Agreeing with Tom below, I would avoid using terms like subclass, decorators, mixin, namespace, (and perhaps even class, instead using objects consistently; this is no place to be pendantic about the distinction between classes and objects; even to the extent of adding a qualifying sentence somewhere like: "In this paper we will use object when we mean object or classes; in software there are important distinctions between the two terms, but they are not relevant in this paper."), and the like and replace them with more generic descriptions, e.g., "extended", or "tools to make defining functions that can use quantities simple to write" }
\inlinecomment{Tom R}{I think saying 'subclass' here is too detailed and is mentioned below in 3.1.1 if needed - 'make it possible to attach units to numpy arrays' or something like that?}

@@

The \texttt{Quantity} object
        is now a subclass of the \texttt{numpy.ndarray} object and therefore
        works well with many of the functions in \texttt{numpy} that support
        array operations. For example, \texttt{Quantity} objects with angular
        units can be passed in to the trigonometric functions implemented in
        \texttt{numpy} (e.g., \texttt{numpy.cos}) and do the right thing: the
        units are internally converted to radians (what the \texttt{numpy}
        trigonometric functions expect) before being passed to \texttt{numpy}.

\subsubsection{Speed improvements}

\inlinecomment{Tom R}{I wonder if we might just want to mention at the top level that performance has been improved in places? This seems a bit too detailed}

The \texttt{Quantity} object is now significantly
        faster than in early versions of \astropypkg. This is in large part due
        to the fact that \texttt{Quantity} subclasses the \texttt{numpy.ndarray}
        object, but even relative to more recent versions of \astropypkg,
        operations with \texttt{Quantity} objects are generally faster. This is
        largely thanks to improvements in the way \texttt{numpy} arrays can be
        subclassed.

@@

The \texttt{astropy.constants} subpackage provides a selection of physical and
astronomical constants as \texttt{Quantity} objects (see
\sectionname~\ref{sec:units}).
A brief description of this package was given in \citep{astropy}.
In version 2.0, the built-in constants have been organized into modules for
specific versions of the constant values.
For example, physical constants have \texttt{codata2014} \citep{codata2014} and
\texttt{codata2010} versions.
Astronomical constants are organized into \texttt{iau2015} and \texttt{iau2012}
modules to indicate their sources (resolutions from the International
Astronomical Union, IAU).
The \texttt{codata2014} and \texttt{iau2015} versions are combined into the
default constant value version: \texttt{astropyconst20}.
For compatibility with \astropypkg version 1.3 and earlier, the \texttt{astropyconst13}
namespace \inlinecomment{Tom R}{jargon alert: 'namespace'} is still available and provides access to the adopted versions of the
constants from earlier versions of \astropypkg.
To use previous versions of the constants as \emph{units} (e.g., solar masses),
the values from the desired namespace have to be imported directly; with version
2.0, \texttt{astropy.units} uses the \texttt{astropyconst20} versions.

@@

New model classes can be created in two ways:
\begin{enumerate}
   \item The simplest way is to use the \texttt{custom\_model} decorator in modeling with a user-defined function that takes the inputs and model parameters as arguments.
   \item It is also possible to subclass the \texttt{Model} base class and create an arbitrarily complex custom model.
\end{enumerate}
